---
title: "The Educational sector"
author: "Adam Abdulrasheed"
date: "2022-09-28"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r main_data}
library(readxl)
df <- read.csv("main_data.csv")
```


```{r pressure, echo=FALSE}
head(df)
```
#Tail
```{r}
tail(df)
```
#dimension of data
```{r}
dim(df)
```
#data structure
```{r}
str(df)
```
```{r}
summary(df)
```
```{r}
attach(df)
```

```{r}
failures.x <- table(failures.x)
```

```{r}
library(likert)
library(gt)
library(gtsummary)

```
```{r}
head(df)
```
#exploring data

```{r}
library(DataExplorer)
```

```{r}
plot_intro(data = df, title = "Visual Representation of Missing Values") 
```







```{r}
v1<-df%>%select(health.x,absences.x)
v1[1:20]<-lapply(v1[1:2], factor,levels=c("Very Bad","Bad","Good","Very Good","Excellent"),ordered=TRUE)
likert(v1)%>%
  plot(type="bar")

```

```{r}
v1<-df%>%select(health.x,failures.x)
v1[1:2]<-lapply(v1[1:2], factor,levels=c("Very Bad","Bad","Good","Very Good","Excellent"),ordered=TRUE)
likert(v1)%>%
  plot(type="bar")
```

# age distribution of the data
```{r}
ag<-table(age)
```

```{r}
barplot(ag)
```


```{r}
df %>%
  group_by(sex) %>%
  tally(sort = T)
```

```{r}
figure1 <- df %>% select(c(sex, failures.x))
figure1 %>% 
  pivot_longer(1:2, names_to = "question", values_to = "response") %>%
  ggplot(aes(x = response)) +
  geom_bar() +
  facet_wrap(vars(question), ncol = 2) +
  labs(x = "Responses", y = "Counts")
```

# Gender with the highest number of failures 
```{r}
df %>%
  group_by(df$sex, df$failures.x)%>%
  tally(sort = T)
```
# gender with most relationships
```{r}
df %>%
  group_by(df$sex, df$studytime.x)%>%
  tally(sort = T)
```



# how parent education affects the failure rates of the children
```{r}
df %>%
  group_by(df$Medu,df$Fedu, df$famsup.x,df$failures.x)%>%
  tally(sort = T)
```

# how family relationship affects the failure of students
```{r}
df %>%
group_by(df$famrel.x,df$failures.x)%>%
tally(sort = T)

```

#family size in relation to the kids taking extra lessons
```{r}

df %>%
  group_by(df$famsize,df$paid.x)%>%
  tally(sort = T)
```

# how the internet access affects the success of the students
```{r}
df %>%
  group_by(df$failures.x,df$internet)%>%
  tally(sort = T)

```
#Failure in response to the desire to go further with the studies
```{r}
df %>%
  group_by(df$failures.x,df$higher.x)%>%
  tally(sort = T)
```
# internet service in relation to the address of the students
```{r}
df %>%
  group_by(df$addRuraless,df$internet)%>%
  tally(sort = T)
```

# how their health status affects their absence from school
```{r}
df %>%
  group_by(df$health.x,df$absences.x,df$failures.x)%>%
  tally(sort = T)
```

```{r}
 library(ggplot2)
```
#failures in relation to gender
```{r}
ggplot(df, aes(x= failures.x, fill= sex)) + geom_bar(position = "dodge")
```
#failure with respect to location
```{r}
ggplot(df, aes(x= failures.x, fill= addRuraless)) + geom_bar(position = "dodge")
```
```{r}
ggplot(df, aes(x= internet, fill= addRuraless)) + geom_bar(position = "stack")
```
```{r}
library(treemapify)
```
```{r}

install.packages("reshape2")
install.packages("plyr")
install.packages("devtools")
```
```{r}
library(devtools)
install_github("wilkox/treemapify")
library(treemapify)
```
```{r}
install.packages('treemapify')
```
```{r}
library(ggplot2)
```



```{r}
ggplot(df, aes(x= school, fill= failures.x)) + geom_bar(position = "stack")
```
```{r}
ggplot(df, aes(x= failures.x, fill= PsTogetherApartTogetherus)) + geom_bar(position = "dodge") #failure with respect to parents living together or apart
```
```{r}
ggplot(df, aes(x= failures.x, fill= guardian.x)) + geom_bar(position = "stack") #failure in relation to guardian
```


correlation between variables

```{r}
library(corrplot)
```
```{r}
corrplot(df, method = "circle")
```
PREDICTIVE ANALYSIS
```{r}
library(rpart)
library(rpart.plot)
```

```{r}
library(caret)
set.seed(123)
inTrain <- createDataPartition(y=df$failures.x, p=0.7, list=FALSE)
```
#partitioning data into testing and training
```{r}
training <- df[inTrain,]
testing <- df[-inTrain,]
```
```{r}
dummies <- dummyVars(failures.x ~ school+ sex + addRuraless + PsTogetherApartTogetherus + health.x + school 
                     + df= training)
dummy_tr <- as.data.frame(predict(dummies, training))
dummy_tr$failures.x <- training$failures.x

modFit <- train(failures.x ~ ., method="treebag", 
                preProc = c("medianImpute"),    
                na.action = na.pass,
                data=dummy_tr)
print(modFit)

```


# Predictive analysis using rpart
```{r}
library(rpart)
library(rpart.plot)
```

```{r}
fit <- rpart(failures.x~., data = training, method = 'class')
print(fit)
rpart.plot(fit)

predict_failures.x<-predict(fit, testing, type = 'class')
print(predict_failures.x)

table_mat <- table(testing$failures.x, predict_failures.x)
table_mat

accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
accuracy_Test
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

